---
title: "Lab 10 Deep Learning Exercises"
author: "Evan Woods"
date: "`r Sys.Date()`"
output: github_document
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
knitr::opts_chunk$set(warning = FALSE)
knitr::opts_chunk$set(message = FALSE)
knitr::opts_chunk$set(fig.width = 6)
knitr::opts_chunk$set(fig.asp = 0.618)
knitr::opts_chunk$set(out.width = "70%")
knitr::opts_chunk$set(fig.align = "center")
knitr::opts_chunk$set(
  comment = ""
)
```

```{r message=FALSE, include=FALSE}
if(!require("MASS")) install.packages("MASS")
if(!require("ISLR2")) install.packages("ISLR2")
if(!require("tidyverse")) install.packages("tidyverse")
if(!require("HH")) install.packages("HH") # VIF
if(!require("e1071")) install.packages("e1071") # naiveBayes
if(!require("class")) install.packages("class") # knn
if(!require("formulaic")) install.packages("formulaic")
if(!require("caTools")) install.packages("caTools")
if(!require("caret")) install.packages("caret")
if(!require("boot")) install.packages("boot")
if(!require("leaps")) install.packages("leaps") # regsubsets
if(!require("glmnet")) install.packages("glmnet") # Ridge and Lasso Regression
if(!require("pls")) install.packages("pls") # Partial Least Squares & Principal Component Regression
if(!require("splines")) install.packages("splines")
if(!require("gam")) install.packages("gam")
if(!require("akima")) install.packages("akima")
if(!require("tree")) install.packages("tree") # Classification and Regression Trees
if(!require("randomForest")) install.packages("randomForest")
if(!require("gbm")) install.packages("gbm") # Boosted Trees
if(!require("BART")) install.packages("BART")
if(!require("reticulate")) install.packages("reticulate") # Use python objects in R
if(!require("ROCR")) install.packages("ROCR")
if(!require("keras")) install.packages("keras") # Install keras for deep learning
if(!require("jpeg")) install.packages("jpeg")
if(!require("imager")) install.packages("imager")
if(!require("survival")) install.packages("survival")

library(survival)
library(imager)
library(keras)
# reticulate::use_condaenv(condaenv = "r-tensorflow")
library(ROCR)
library(reticulate)
library(BART)
library(gbm)
library(randomForest)
library(tree)
library(akima)
library(gam)
library(splines)
library(glmnet)
library(pls)
library(leaps)
library(formulaic)
library(class)
library(e1071)
library(HH)
library(MASS)
library(ISLR2)
library(tidyverse)
library(caTools)
library(caret)
library(boot)
library(jpeg)
```

```{r output=FALSE, results = 'hide', message=FALSE}
# keras::install_keras(method = "conda", python_version = "3.10")
```

```{r include=FALSE}
custom_darkblue = "#1A0875"
custom_lightblue = "#34ABEB"
custom_red = "#a60808"
```

```{r include=FALSE}
f_print <- function(string){
  cat(str_wrap(string = string, width = 80, indent = 0, exdent = 0, whitespace_only = TRUE))
}
```

## Applied
### Question 10:
This exercise focuses on the brain tumor data, which is included in the ISLR2 R library.

* **Question 10-a**: Plot the Kaplan-Meier survival curve with ±1 standard error bands, using the survfit() function in the survival package.
  * **Answer**:
```{r}
attach(BrainCancer)
fit.surv <- survfit(Surv(time, status) ~ 1)
summary(fit.surv)
plot(fit.surv)
```
* **Question 10-b**: Draw a bootstrap sample of size n = 88 from the pairs (y, δ), and compute the resulting Kaplan-Meier survival curve. Repeat this process B = 200 times. Use the results to obtain an estimate of the standard error of the Kaplan-Meier survival curve at each timepoint. Compare this to the standard errors obtained in (a). 
  * **Answer**: 
```{r}
boot <- sample(c(time, status), size = 88)
fit.surv <- survfit(Surv(boot) ~ 1)
plot(fit.surv)
boot <- list()
```

```{r}
boot[[1]] <- sample(c(time, status), size = 88)
fit.surv <- survfit(Surv(boot[[1]]) ~ 1)
fit.surv$std.err
plot(fit.surv)
```


```{r}
boot_list <- boot[[1]]
for (i in seq(1, 200)) {
  boot[[i]] <- sample(c(time, status), size = 88)
}
```


