---
title: "Lab 5 Cross-Validation and the Bootstrap Exercises"
author: "Evan Woods"
date: "`r Sys.Date()`"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(fig.width = 6)
knitr::opts_chunk$set(fig.asp = 0.618)
knitr::opts_chunk$set(out.width = "70%")
knitr::opts_chunk$set(fig.align = "center")
knitr::opts_chunk$set(
  comment = ""
)
```

```{r message=FALSE, include=FALSE}
if(!require("MASS")) install.packages("MASS")
if(!require("ISLR2")) install.packages("ISLR2")
if(!require("tidyverse")) install.packages("tidyverse")
if(!require("HH")) install.packages("HH") # VIF
if(!require("e1071")) install.packages("e1071") # naiveBayes
if(!require("class")) install.packages("class") # knn
if(!require("formulaic")) install.packages("formulaic")
if(!require("caTools")) install.packages("caTools")
if(!require("caret")) install.packages("caret")
if(!require("boot")) install.packages("boot")

library(boot)
library(formulaic)
library(class)
library(e1071)
library(HH)
library(MASS)
library(ISLR2)
library(tidyverse)
library(caTools)
library(caret)
```

```{r include=FALSE}
LoadLibraries <- function() {
if(!require("MASS")) install.packages("MASS")
if(!require("ISLR2")) install.packages("ISLR2")
if(!require("tidyverse")) install.packages("tidyverse")
if(!require("HH")) install.packages("HH") # VIF
if(!require("e1071")) install.packages("e1071") # naiveBayes
if(!require("class")) install.packages("class") # knn
if(!require("formulaic")) install.packages("formulaic")
if(!require("caTools")) install.packages("caTools")
if(!require("caret")) install.packages("caret")

library(caret)
library(formulaic)
library(class)
library(e1071)
library(HH)
library(MASS)
library(ISLR2)
library(tidyverse)
library(caTools)
  print("Libraries have been loaded!")
}
```

```{r include=FALSE}
custom_darkblue = "#1A0875"
custom_lightblue = "#34ABEB"
custom_red = "#a60808"
```


```{r include=FALSE}
f_print <- function(string){
  cat(str_wrap(string = string, width = 80, indent = 0, exdent = 0, whitespace_only = TRUE))
}
```

```{r include=FALSE}
set.seed(42)
attach(Default)
```


## Applied

### Question 5:
In Chapter 4, we used logistic regression to predict the probability of
default using income and balance on the Default data set. We will
now estimate the test error of this logistic regression model using the
validation set approach. Do not forget to set a random seed before
beginning your analysis.


* **Question-5a**: Fit a logistic regression model that uses income and balance to predict default.
  * **Answer**:
```{r}
glm.fit <- glm(default ~ income + balance, data = Default, family = "binomial")
```


* **Question-5b**: Using the validation set approach, estimate the test error of this model. In order to do this, you must perform the following steps:
    i. Split the sample set into a training set and a validation set. 
    ii. Fit a multiple logistic regression model using only the training observations.
    iii. Obtain a prediction of default status for each individual in
the validation set by computing the posterior probability of
default for that individual, and classifying the individual to
the default category if the posterior probability is greater
than 0.5.
    iv. Compute the validation set error, which is the fraction of
the observations in the validation set that are misclassified.
  * **Answer**:
```{r}
# Split the sample set into a training set and a validation set.
train <- sample(nrow(Default), (nrow(Default) * .8))
```

```{r}
# Fit a multiple logistic regression model using only the training observations.
glm.fit <- glm(default ~ income + balance, data = Default, family = "binomial", subset = train)
```

```{r}
# Obtain a prediction of default status for each individual in the validation set by computing the posterior probability of default for that individual, and classifying the individual to the default category if the posterior probability is greater than 0.5. 
glm.prob <- predict(glm.fit, Default[-train, 2:4], type = "response")
glm.pred <- rep("No", length(Default$default))
glm.pred[glm.prob > .5] <- "Yes"
```

```{r echo=FALSE}
# Compute the validation set error, which is the fraction of the observations in the validation set that are misclassified. 
# as.factor(Default$default[-train])
f_print(sprintf("The validation set error is: %0.3f%%.", (mean(glm.pred != Default$default[-train]) * 100)))
```

* **Question 5-c**: Repeat the process in the last question three times, using three different splits of the observations into a training set and a validation set. Comment on the results obtained. 
  * **Answer**: 
```{r warning=FALSE, echo=FALSE}
set.seed(42)
split <- c(.6, .7, .9)
for (i in rep(1:3)){
  train <- sample(nrow(Default), (nrow(Default) * split[[i]]))
  glm.fit <- glm(default ~ income + balance, data = Default, family = "binomial", subset = train) 
  glm.prob <- predict(glm.fit, Default[-train, 2:4], type = "response")
  glm.pred <- rep("No", length(Default$default))
  glm.pred[glm.prob > .5] <- "Yes"
  
  f_print(sprintf("The validation set error is: %0.3f%% for the split of %0.2f%%.", (mean(glm.pred != Default$default[-train]) * 100), split[[i]]))
  cat("\n")
}
```

```{r echo=FALSE}
f_print(sprintf("The validation set error is greatest for the 60%% split. The validation set error is smallest when using a 70%% split."))
```


* **Question 5-d**: Now consider a logistic regression model that predicts the probability of default using income, balance, and a dummy variable for student. Estimate the test error for this model using the validation set approach. Comment on whether or not including a dummy variable for student leads to a reduction in the test error rate. 
  * **Answer**:
```{r echo=FALSE}
set.seed(42)

default_yes <- Default %>% filter(student == "Yes") %>% mutate(d_student = 1) %>% select(d_student, everything())
default_no <- Default %>% filter(student == "No") %>% mutate(d_student = 0) %>% select(d_student, everything())
default <- add_row(default_yes, default_no)
default <- default %>% select(default, d_student, balance, income, student)
default_train <- default %>% mutate(train = sample.split(default, .8))
train <- default_train %>% filter(train == TRUE) %>% select(everything(), -train)
test <- default_train %>% filter(train == FALSE) %>% select(everything(), -train)

glm.fit <- glm(default ~ income + balance + d_student, data = train, family = "binomial")
glm.prob <- predict(glm.fit, test[2:4], type = "response")
glm.pred <- rep("No", length(test$default))
glm.pred[glm.prob > 0.5] <- "Yes"
f_print(sprintf("The test error using the validation set approach with a dummy variable is: %0.2f%%.",(mean(glm.pred != test$default) * 100)))
cat("\n\n")

glm.fit <- glm(default ~ income + balance + student, data = train, family = "binomial")
glm.prob <- predict(glm.fit, test[3:5], type = "response")
glm.pred <- rep("No", length(test$default))
glm.pred[glm.prob > 0.5] <- "Yes"
f_print(sprintf("The test error using the validation set approach with a student variable as a factor is the same at: %0.2f%%.",(mean(glm.pred != test$default) * 100)))
cat("\n\n")

glm.fit <- glm(default ~ income + balance, data = train, family = "binomial")
glm.prob <- predict(glm.fit, test[3:4], type = "response")
glm.pred <- rep("No", length(test$default))
glm.pred[glm.prob > 0.5] <- "Yes"
f_print(sprintf("The test error using the validation set approach without a dummy student variable results in decreased test error observed at: %0.2f%%.",(mean(glm.pred != test$default) * 100)))
cat("\n\n")

f_print(sprintf("The inclusion of a dummy variable for the student predictor lowers the test error."))
```

### Question 6: We continue to consider the use of a logistic regression model to
predict the probability of default using income and balance on the
Default data set. In particular, we will now compute estimates for
the standard errors of the income and balance logistic regression coef-
ficients in two different ways: (1) using the bootstrap, and (2) using
the standard formula for computing the standard errors in the glm()
function. Do not forget to set a random seed before beginning your
analysis.

* **Question 6-a**: Using the summary() and the glm() functions, determine the standard errors for the coefficients associated with income and balance in a multiple logistic regression model that uses both predictors. 
  * **Answer**:
  
```{r echo=FALSE}
set.seed(42)
summary_fn <- summary(glm(default ~ income + balance, data = train, family = "binomial"))$coefficients
```
* **Question 6-b**: Write a function, boot.fn(), that takes as input the Default data
set as well as an index of the observations, and that outputs
the coefficient estimates for income and balance in the multiple
logistic regression model.
  * **Answer**:
```{r}
boot.fn <- function(data, index) {
  coef(glm(default ~ income + balance, data = data, subset = index, family = "binomial"))
}
```

* **Question 6-c**: Use the boot() function together with your boot.fn() function to
estimate the standard errors of the logistic regression coefficients
for income and balance.
  * **Answer**:
```{r echo=FALSE}
set.seed(42)
bootstrap_fn <- boot(Default, boot.fn, R = 1000)
bootstrap_fn
```




* **Question 6-d**: Comment on the estimated standard errors obtained using the
glm() function and using your bootstrap function.
  * **Answer**:
```{r echo=FALSE}
f_print(sprintf("The standard errors obtained for the coefficients while using the summary function are: %0.3e and %0.3e respectively.", summary_fn[2,2], summary_fn[3,2]))
f_print(sprintf("The standard errors obtained for the coefficients while using the bootstrapped datasets are lower for both coefficients."))
```

### Question 7: In Sections 5.3.2 and 5.3.3, we saw that the cv.glm() function can be
used in order to compute the LOOCV test error estimate. Alterna-
tively, one could compute those quantities using just the glm() and
predict.glm() functions, and a for loop. You will now take this ap-
proach in order to compute the LOOCV error for a simple logistic
regression model on the Weekly data set. Recall that in the context
of classification problems, the LOOCV error is given in (5.4).

```{r include=FALSE}
weekly <- na.omit(Weekly)
attach(weekly)
```

* **Question 7-a**: Fit a logistic regression model that predicts Direction using Lag1
and Lag2.
  * **Answer**:
```{r echo=FALSE}
glm(Direction ~ Lag1 + Lag2, data = Direction, family = "binomial")
```

* **Question 7-b**: Fit a logistic regression model that predicts Direction using Lag1
and Lag2 using all but the first observation.
  * **Answer**:
```{r}
glm.fit_loo <- glm(Direction ~ Lag1 + Lag2, data = weekly[-1,], family = "binomial")
```

* **Question 7-c**: Use the model from the previous question to predict the direction of the first obser-
vation. You can do this by predicting that the first observation
will go up if P (Direction = "Up"|Lag1, Lag2) > 0.5. Was this
observation correctly classified?
  * **Answer**:
```{r echo=FALSE}
glm.fit_loo_pred <- rep("Down", 1)
glm.fit_loo_prob <- predict(glm.fit_loo, weekly[1,], type = "response")
glm.fit_loo_pred[glm.fit_loo_prob > 0.5] <- "Up"
f_print(sprintf("The statement that this prediction was correctly classified is: %s", (glm.fit_loo_pred == weekly$Direction[1])))
```


* **Question 7-d**: Write a for loop from i = 1 to i = n, where n is the number of
observations in the data set, that performs each of the following
steps:
  i. Fit a logistic regression model using all but the ith obser-
vation to predict Direction using Lag1 and Lag2.
  ii. Compute the posterior probability of the market moving up
for the ith observation.
  iii. Use the posterior probability for the ith observation in order
to predict whether or not the market moves up.
  iv. Determine whether or not an error was made in predicting
the direction for the ith observation. If an error was made,
then indicate this as a 1, and otherwise indicate it as a 0.
  * **Answer**:
```{r}
loocv <- function(n) {
    truthiness <- c()
    for (i in seq(1, n)) {
      glm.fit_loo <- glm(Direction ~ Lag1 + Lag2, data = weekly[-i,], family = "binomial")      
      glm.fit_loo_pred <- rep("Down", 1)
      glm.fit_loo_prob <- predict(glm.fit_loo, weekly[i,], type = "response")
      glm.fit_loo_pred[glm.fit_loo_prob > 0.5] <- "Up"
      
      if(glm.fit_loo_pred != weekly$Direction[i]){
        truthiness <- c(truthiness, 1)
      } else {
        truthiness <- c(truthiness, 0)
      }
    }
    return(truthiness)
  }
```
* **Question 7-e**: Take the average of the n numbers obtained in (d)iv in order to
obtain the LOOCV estimate for the test error. Comment on the
results.
  * **Answer**: 
```{r echo=FALSE}
n = 1000
f_print(sprintf("The average error of the %0.0f numbers obtained in the custom leave-one-out-cross-validation function is: %0.2f%%. This appears to be an appropriate level of error for this dataset.", n, mean(loocv(n)) * 100))
```

### Question 8: We will now perform cross-validation on a simulated data set. 

* **Question 8-a**: Generate the simulated data set. In this dataset, what is _n_ and what is _p_?
  * **Answer**:
```{r}
set.seed(1)
x <- rnorm(100)
y <- x - 2 * x^2 + rnorm(100)
```

```{r echo=FALSE}
  f_print(sprintf("In this dataset, there are 100 observations, n, & 2 predictors, p."))
```

* **Question 8-b**: Create a scatterplot of _X_ against _Y_. Comment on what you find. 
  * **Answer**:
```{r echo=FALSE}
ggplot() +
  geom_point(aes(x, y)) +
  labs(title = expression('ƒ(x) = x - 2 * x'^{2} * + 'ε'), y = "ƒ(x)", x = 'x') +
  theme_linedraw()
f_print(sprintf("The resulting function is a quadratic with an upside-down parabolic curve. There is observable noise in the scatterplot."))
```


* **Question 8-c**: Set a random seed, and then compute the LOOCV errors that
result from fitting the following four models using least squares.
  i. Y = β~0~ + β~1~X + ε
  ii. Y = β~0~ + β~1~X + β~2~X^2^ + ε
  iii. Y = β~0~ + β~1~X + β~2~X^2^ + β~3~X^3^ + ε
  iv. Y = β~0~ + β~1~X + β~2~X^2^ + β~3~X^3^ + β~4~X^4^ + ε
  * **Answer**:

```{r}
set.seed(1)
x <- rnorm(100)
y <- x - 2 * x^2 + rnorm(100)
df <- tibble(x,y)
# i.
glm.fit_i <- glm(y ~ x, data = df)
cv.error.i <- cv.glm(df, glm.fit_i)$delta[1]

# ii.
glm.fit_ii <- glm(y ~ x + I(x^2), data = df)
cv.error.ii <- cv.glm(df, glm.fit_ii)$delta[1]

# iii.
glm.fit_iii <- glm(y ~ x + I(x^2) + I(x^3), data = df)
cv.error.iii <- cv.glm(df, glm.fit_iii)$delta[1]

# iv.
glm.fit_iv <- glm(y ~ x + I(x^2) + I(x^3) + I(x^4), data = df)
cv.error.iv <- cv.glm(df, glm.fit_iv)$delta[1]
f_print(sprintf("The LOOCV errors that result from fitting the four models are as follows: %0.03f%%, %0.03f%%, %0.03f%%, %0.03f%%", cv.error.i, cv.error.ii, cv.error.iii, cv.error.iv))
```


* **Question 8-d**: Repeat the previous question using another random seed, and report your results.
Are your results the same as what you got in the previous question? Why?
  * **Answer**:

```{r}
set.seed(42)
x <- rnorm(100)
y <- x - 2 * x^2 + rnorm(100)
df <- tibble(x,y)
# i.
glm.fit_i <- glm(y ~ x, data = df)
cv.error.i <- cv.glm(df, glm.fit_i)

# ii.
glm.fit_ii <- glm(y ~ x + I(x^2), data = df)
cv.error.ii <- cv.glm(df, glm.fit_ii)$delta[1]

# iii.
glm.fit_iii <- glm(y ~ x + I(x^2) + I(x^3), data = df)
cv.error.iii <- cv.glm(df, glm.fit_iii)$delta[1]

# iv.
glm.fit_iv <- glm(y ~ x + I(x^2) + I(x^3) + I(x^4), data = df)
cv.error.iv <- cv.glm(df, glm.fit_iv)$delta[1]
f_print(sprintf("With a new seed, the LOOCV errors that result from fitting the four models are as follows: %0.03f%%, %0.03f%%, %0.03f%%, %0.03f%%", cv.error.i$delta[1], cv.error.ii, cv.error.iii, cv.error.iv))
```

```{r echo=FALSE}
f_print(sprintf("The results differ from the previous question. With a new seed, new random number are generated. The variability in the error term is irreducible error that persists between models. It is simulated when new data is pseudo-randomly generated by setting a new seed, this irreducible error is observable in the differing values of the test errors the models exhibited.")) 
```
* **Question 8-e**: Which of the models had the smallest LOOCV error? Is this what was expected? Explain your answer.
  * **Answer**:
```{r echo=FALSE}
f_print(sprintf("The model whose formula was composed of a highest degree of 2 had the smallest error as expected. This predictors of this model match that of the true function of f which is itself comprised of predictors x and x squared as well."))
```

* **Question 8-f**: Comment on the statistical significance of the coefficient esti-
mates that results from fitting each of the models using
least squares. Do these results agree with the conclusions drawn
based on the cross-validation results?
  * **Answer**:

```{r echo=FALSE}
summary(glm.fit_i)
summary(glm.fit_ii)
summary(glm.fit_iii)
summary(glm.fit_iv)
f_print(sprintf("In all models, when present, the predictors x and x-squared were found to be statistically significant predictors of the response, y. These results agree with the conclusions drawn based on the cross-validation results where the test error was least significant when these two variables were exclusively present. Models which included additional predictors that did not add value to the model increased test error. These same models were also observed to not have statistical significance with respect to the coefficients of the same predictors as shown from the model summaries."))
```
### Question 9:
We will now consider the Boston housing dataset, from the ISLR2 library.

* **Question 9-a**: Based on this data set, provide an estimate for the population
mean of medv. Call this estimate ˆμ.
  * **Answer**:
```{r include = FALSE}
boston <- na.omit(Boston)
attach(boston)
```

```{r}
μ_ˆ <- mean(boston$medv)
```

```{r echo=FALSE}
f_print(sprintf("The estimate of the population mean of medv is: %0.3f.",μ_ˆ))
```


* **Question 9-b**: Provide an estimate of the standard error of ˆμ. Interpret this
result.
  * **Answer**:
```{r echo=FALSE}
provided_se <- sd(boston$medv)/sqrt(length(boston$medv))
f_print(sprintf("The estimated standard error, ˆμ, is: %0.3f.",provided_se))
```

* **Question 9-c**: Now estimate the standard error of ˆμ using the bootstrap. How
does this compare to your answer from the previous question?
  * **Answer**:
```{r echo=FALSE}
R = 1000
set.seed(42)
boot.fn <- function(data, index) {
  mean(data[index])
}

boot_mean <- boot(boston$medv, boot.fn, R = 1000)

boot_mean
se <- 0.4009216

f_print(sprintf("The standard error of %0.0f bootstrapped sample means is: %0.3f. The estimated standard error using the bootstrap is approximately the same as the estimated standard error with a value of: %0.3f.", R, se, provided_se))
```


* **Question 9-d**: Based on your bootstrap estimate from the previous question, provide a 95 % con-
fidence interval for the mean of medv. Compare it to the results
obtained using t.test(Boston$medv).
  * **Answer**:
```{r}
conf_int <- c(μ_ˆ- 2*se, μ_ˆ+ 2*se)
conf_int
t_test <- t.test(boston$medv)
t_test
```


```{r echo=FALSE}
f_print(sprintf("The calculated confidence interval ranges from %0.3f to %0.3f. These values are very similar to the 95%% confidence interval as calculated by the t test.", conf_int[1], conf_int[2]))
```
* **Question 9-e**: Based on this data set, provide an estimate, ˆμmed, for the median
value of medv in the population.
  * **Answer**:
```{r echo=FALSE}
μ_ˆ_med <- median(boston$medv)
f_print(sprintf("The estimate of the median of the population mean of the boston medv variable is: %0.2f.", μ_ˆ_med))
```

* **Question 9-f**: We now would like to estimate the standard error of ˆμmed. Unfor-
tunately, there is no simple formula for computing the standard
error of the median. Instead, estimate the standard error of the
median using the bootstrap. Comment on your findings.
  * **Answer**:
```{r echo=FALSE}
R = 1000
set.seed(42)
boot.fn <- function(data, index) {
  median(data[index])
}
boot_median <- boot(boston$medv, boot.fn, R = 1000)
boot_median
boot_median_se <- 0.3661785
f_print(sprintf("The standard error of %0.0f bootstrapped sample medians is: %0.3f.", R, boot_median_se))
```

* **Question 9-g**: Based on this data set, provide an estimate for the tenth per-
centile of medv in Boston census tracts. Call this quantity ˆμ0.1.
(You can use the quantile() function.)
  * **Answer**:
```{r echo=FALSE}
μ_ˆ_0_1 <- quantile(boston$medv, probs = c(.1))
# μ_ˆ_0_1
```


```{r echo=FALSE}
f_print(sprintf("The estimate for the tenth percentile of medv in Boston census tracts, ˆμ0.1, is: %0.2f.", μ_ˆ_0_1))
```

* **Question 9-h**: 
  * **Answer**:
```{r echo=FALSE}
set.seed(42)
boot.fn <- function(data, index) {
  quantile(data[index], probs = c(.1))
}

boot(boston$medv, boot.fn, R = 1000)
f_print(sprintf("The bootstrapped estimate of the tenth percentile of medv in the Boston census tracts is identical to the estimate created using the quantile function. The standard error is 0.494."))
```
  

